---
title: "ST 502 HW3 Question 5"
author: "Halid Kopanski"
date: "2/2/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Creation and Set Up

```{r}
library(ggplot2)

par(mfrow = c(2 , 2))

set.seed(1000)

N = 1000

df_data <- data.frame(cbind(rnorm(N, 0, 1),
                            rnorm(N, 0, 1),
                            rnorm(N, 0, 1))
                      )

x_names <- c("X1", "X2", "X3")

df_z <- data.frame(cbind(
  
  df_data[,1] + 2*df_data[,2] + 3*df_data[,3],
    
  df_data[,1]^2 + df_data[,2]^2 + df_data[,3]^2,
    
  0.5*(df_data[,1] - df_data[,3])^2,
    
  (2*df_data[,3]^2) / (df_data[,1]^2 + df_data[,2]^2)
                        )
  
                   )

z_names <- c("Za", "Zb", "Zc", "Zd")

colnames(df_z) <- z_names

chart_colors <- c("#003f5c", "#2f4b7c", "#665191", "#a05195",
                  "#d45087", "#f95d6a", "#ff7c43", "#ffa600")
```
\newpage
## Plots of X random variables

The mean and variance for each variable is quite close to 0 and 1, respectively. The reason for the values not be exact is due to noise in the generated data. Generally speaking, all three random variables follow a normal like distribution. Which was expected.

```{r}
for (i in 1:3){
  stmnt <- sprintf("The mean of %s is %.4f and the variance is %.4f", 
                   x_names[i], 
                   mean(df_data[ ,i]), 
                   var(df_data[ ,i])
                   )
  
  print(stmnt)
}

for (i in 1:3){
    
  stmnt <- sprintf("Mean: %.4f \n Variance: %.4f",
                   mean(df_data[ ,i]), 
                   var(df_data[ ,i])
                   )

  dplot <- ggplot(df_data, aes( x = df_data[, i])) + 
           geom_density(fill = chart_colors[i + 4], alpha = 0.7) +
           ggtitle(paste0("X", i)) + xlab("") + ylab("") + 
           theme(plot.title = element_text(hjust = 0.5)) +
           geom_text(x = 0, y = 0.2, label = stmnt) + facet_grid()
  
  print(dplot)
}
```

\newpage
## Plots of Z Variables

Plotting the Z variables, we can see that they form a normal, chi squared, gamma, and F distributions as expected 
from the equations used to construct each Z variable. 

The mean and variance of each Z variable is quite close to the computed ones found in Problem 4. The differences are
most likely due the noise observed in the X variable data. When plotting the Z1, Z2, and Z3 variables against data generated directly from the suspected distribution. This can be observed in the charts where each Z variable is overlaid on the equivalent distribution. They roughly matched the peak and spread.

Z4 was difficult to ascertain since the Z variable and the distribution were both extreme. Zooming into the plot helped view the comparison better.

```{r}
set.seed(1000)

df_distros <- data.frame(cbind(rnorm(N, 0, 14^0.5), 
                               rchisq(N, 3), 
                               rchisq(N, 1), 
                               rf(N, 1, 2)))

for (i in 1:4){
  stmnt <- sprintf("Mean: %.4f Variance: %.4f",
                   mean(df_z[ ,i]), 
                   var(df_z[ ,i])
                   )
  dplot <- ggplot() + 
           geom_density(aes( x = df_z[, i]), fill = chart_colors[i], alpha = 0.7) +
           geom_density(aes( x = df_distros[, i]), fill = chart_colors[i + 4], alpha = 0.7) +
           ggtitle(label = paste0("Z", i), subtitle = stmnt) + xlab("") + ylab("") + 
           theme(plot.title = element_text(hjust = 0.5))
  
  print(dplot)
}
```

\newpage

The following chart zooms into the F distribution and reveals that the two variables match quite closely:

```{r, warning = FALSE, message = FALSE}

ggplot() + 
           geom_density(aes( x = df_z[, 4]), fill = chart_colors[4], alpha = 0.7) +
           geom_density(aes( x = df_distros[, 4]), fill = chart_colors[4], alpha = 0.7) +
           ggtitle(label = paste0("Z", i), subtitle = stmnt) + xlab("") + ylab("") + 
           theme(plot.title = element_text(hjust = 0.5)) + xlim(0, 15)

```


```{r}

mean_sd_info <- rep(0, 4)

for (i in 1:4){
    x_info <- c(mean(df_z[ ,i]), var(df_z[ ,i]), mean(df_distros[ ,i]), var(df_distros[ ,i]))
    mean_sd_info <- rbind(mean_sd_info, x_info)
}

mean_sd_info <- mean_sd_info[2:5, ]

colnames(mean_sd_info) <- c("Z Mean", "Z Var", "Distro Mean", "Distro Var")
rownames(mean_sd_info) <- c("Z1", "Z2", "Z3", "Z4")

knitr::kable(mean_sd_info)
```


As can be seen in the above table, the Z variables and their respective distribution mean and variance were quite close but not exactly the same. This is due to how the data was generated. All generated data sets used in this document displayed noise. This noise caused noticeable changes in the density plots. Overall, the Z variable and the respective distribution matched quite closely, verifying the results in the question 4.